\documentclass{../notes}
\title{Flipping for quick profit or quick loss}
\begin{document}
\maketitle

\section{Setup}
\begin{itemize}
  \item Presence of a good impulse setting up further continuation (easily 2-3x larger volume)
  \item Absence of momentum divergence. (Even if within one area of high-low? Jagged high?)
  \item Absence of buying or selling climax.
  \item Pullback is above half of trend-leg high. Start counting trend-leg from just when fast-line breaks above 0.
  \item Market is in a clear uptrend with visible long-term higher lows and higher highs.
  \item ????: Make sure no high-volume downward momentum in past rallies (big spike down--even if rebounded immediately)
  \item ???: New high must be equal or greater slope compared to rest of trend
  \item ?????: No 3 peaked MACD fast-line spike reminicent of 3-pushes
\end{itemize}

\section{Trigger}
Buy when MACD fast line crosses slow line when all of setup is met.
\begin{itemize}
  \item ???? Buy on red candle near end of hour?
\end{itemize}

\section{Stop}
Highest of:
\begin{itemize}
  \item Bottom 25\% of trend-leg
  \item 50\% lower than buy price
\end{itemize}
\section{Profit Target}
\begin{itemize}
  \item 1x and 2x risk multiple. Sell 50/50 each. If 1x risk multiple is hit move stop-loss to buy-price.
\end{itemize}
\section{Comments}


\section{Setup-Trigger backtest game}
\begin{itemize}
  \item Add • if criteria was met and led to a successful trade
  \item Add -- if criteria was not met and trade was not successful
  \item Remove • if criteria was not met and trade was successful
  \item If reach line, then it is done ---------------------------------------->
\end{itemize}
\begin{itemize}
  \item Presence of a good impulse setting up further continuation (easily 2-3x larger volume min) \\
  \item Absence of momentum divergence. (Even if within one area of high-low? Jagged high?)
  \item Absence of buying or selling climax.
  \item Market is in a clear uptrend with visible long-term higher lows and higher highs

  \item Pullback closeis above half of trend-leg high. Start counting trend-leg from just when fast-line breaks above 0
        ---•--•--•--•••----•
        X
  \item Pullback low is above half of trend-leg high. Start counting trend-leg from just when fast-line breaks above 0
        -----•--•----•••----•
        X
  \item ????: Make sure no visible super high-volume downward momentum in past rallies (big spike down--even if rebounded immediately) for past few days \\
        •--•••••--•--
  \item ???: New high must be equal or greater slope compared to rest of trend\\
        •••••••••••
  \item ?????: No 3 peaked MACD fast-line spike reminicent of 3-pushes\\
        ••••••--•••••
  \item ???? Buy on red candle near end of hour?\\
        •----••••••••
  \item Impulse buy made of many medium sized green candles (3+)
        •--•••----••••
  \item No medium size sell-off between previous push and this push (sell-off approx size of previous push buy)
        •••
  \item Not concave up and crashing down
        •--••--••----•
  \item No streak of fans for past month
        --••••
  \item Has at least a strong green candle with volume easily 5x
        ••••
  \item INJ: OOO




\end{itemize}








\end{document}
